A Hybrid Neutronics Method for Time-Dependent Control Rod Modeling in Molten Salt Reactors

### 2024-11-07 msre.tex
Rename files

\subsection{2-D Neutronics Model Setup}

Figures \ref{fig:1/4-geom} and \ref{fig:full-geom} show the 2-D quarter-core and full-core
\gls{MSRE} models investigated. I adapted the models from the horizontal cross section of the base
\gls{MSRE} benchmark model developed by \cite{fratoni_molten_2020} with the following differences
in the 2-D \gls{MSRE} model for this work:

\begin{itemize}
  \item Ignored thermal expansion of all reactor components.
  \item Modeled the control rods as perfectly annular cylinders instead of beaded control elements
    strung on flexible steel cables. Ignored the inconel shells encasing the control elements and
    the steel helical cables linking the control elements together.
  \item Set channel pitch, width, and length to 2, 0.2, and 1.2 inches, respectively.
  \item Homogenized the INOR, graphite, and molten salt regions in the sample basket.
  \item Modified the outer edge of the core graphite to be circular with no jagged edges.
  \item Replaced partial channels with full-size channels near the edge of the core graphite.
  \item Ignored the thermal insulation and shield structures outside the core.
\end{itemize}

@@

\subsection{2-D Neutronics Numerical Results} \label{sec:2d-nts-results}

\subsubsection{2-D quarter-core results}

@@

TBD

%\subsection{3-D Model Setup}

%\subsection{3-D Numerical Results}

@@

TBD

%The hybrid method does not alter the $k_\text{eff}$ and flux estimates relative to the neutron
%diffusion method in cases where the neutron diffusion method provides

>>>>

\subsection{2-D Neutronics Model Setup} \label{sec:2d-model-setup}

Figures \ref{fig:1/4-geom} and \ref{fig:full-geom} show the 2-D quarter-core and full-core
\gls{MSRE} models investigated. I adapted the models from the horizontal cross section of the
\gls{MSRE} numerical benchmark model \cite{fratoni_molten_2020} with the
following differences in the 2-D \gls{MSRE} model for this work:

\begin{itemize}
  \item Ignored thermal expansion of all reactor components.
  \item Modeled the control rods as perfectly annular cylinders instead of beaded control elements
    strung on flexible steel cables. Ignored the inconel shells encasing the control elements and
    the steel helical cables linking the control elements together.
  \item Set channel pitch, width, and length to 2, 0.2, and 1.2 inches, respectively.
  \item Homogenized the INOR-8, graphite, and molten salt regions in the sample basket.
  \item Modified the outer edge of the core graphite to be circular with no jagged edges.
  \item Replaced partial channels with full-size channels near the edge of the core graphite.
  \item Ignored the thermal insulation and shield structures outside the core.
\end{itemize}

@@

\subsection{2-D Neutronics Numerical Results \& Discussion} \label{sec:2d-nts-results}

\subsubsection{2-D Quarter-Core}

@@

\subsection{3-D Neutronics Model Setup} \label{sec:3d-model-setup}

Figure \ref{fig:msre-picture} shows the vertical cross section of the actual \gls{MSRE} vessel.
The actual design has rounded upper and lower plena to help direct salt flow through the system.
This work uses a modified 3-D numerical \gls{MSRE} model as shown by the vertical cross section
in Figure \ref{fig:msre-geom-vert}. The model excludes the toroidal flow distributor near the top
of the vessel and reshapes the plena into right cylinders. It has the same horizontal cross section
as the 2-D model (Figure \ref{fig:full-geom}) throughout most of the salt-graphite lattice. In
addition to modifications in the 2-D model listed in Section \ref{sec:2d-model-setup}, the
3-D model includes the following modifications:

\begin{itemize}
  \item Removed the toroidal flow distributor near the top of the vessel.
  \item Reshaped the upper and lower plena into right cylinders while keeping the salt volume
    constant.
  \item Flattened the conical tops of the graphite lattice blocks.
  \item Regularized the complicated graphite lattice structure at the bottom.
  \item Set a fixed outer vessel thickness of 2.54 cm.
  \item Homogenized the lower plenum by 90.8\% salt- 9.2\% INOR-8 (vessel) by volume.
  \item Removed the reactor access nozzle and the various components within it.
  \item Added a graphite block at the bottom of the control rod thimbles.
\end{itemize}

\begin{figure}[p]
  \centering
  \includegraphics[width=0.97\columnwidth]{msre-picture}
  \caption{Vertical cross section of the actual \gls{MSRE} vessel. It has rounded upper and lower
  plena, and a toroidal flow distributor that distributes the salt inflow evenly as it enters the
  core through holes on the inside vessel surface of the distributor.}
  \label{fig:msre-picture}
\end{figure}

\begin{figure}[p]
  \centering
  \includegraphics[width=\columnwidth]{msre-geom-vert}
  \caption{Vertical cross section of the 3-D numerical \gls{MSRE} model offset by 5.08 cm to show
  the control rod thimble and homogenized sample basket. The control rod is in its fully inserted
  position. The homogenized lower plenum consists of 90.8\% salt and 9.2\% INOR-8 alloy by volume.}
  \label{fig:msre-geom-vert}
\end{figure}

Table \ref{table:reactor-dimensions} lists relevant dimensions of the 3-D \gls{MSRE} model.
All material compositions remain
the same as the 2-D models and the \gls{MSRE} benchmark specifications \cite{fratoni_molten_2020}.
All 3-D neutronics simulations in this section used the molten salt composition with 65.25 kg
$^{235}$U loaded (Table \ref{table:2d-salt-composition}), except the simulations for temperature
reactivity coefficient calculations which had a higher $^{235}$U loading of 67.00 kg (Table
\ref{table:67-salt-composition}).

\begin{table}[htb]
  \centering
  \caption{List of reactor dimensions.}
  \begin{tabular}{l l}
    \toprule
    Height [cm] & 215 \\
    \midrule
    TODO: Populate table & \\
    \bottomrule
  \end{tabular}
  \label{table:reactor-dimensions}
\end{table}

\begin{table}[htb]
  \small
  \centering
  \setlength\tabcolsep{4pt}
  \caption{\gls{MSRE} molten salt composition when the $^{235}$U loading was at 65.25 kg.}
  \begin{tabular}{l S S S S S S S}
    \toprule
    Element & {Li} & {Be} & {Zr} & {Hf} & {$^{234}$U} & {$^{235}$U} & {$^{236}$U} \\
    \midrule
    Composition [kg] & 507.42 & 293.96 & 513.97 & 0.0029 & 0.68 & 67.00 & 0.28 \\
    \bottomrule
  \end{tabular}
  \begin{tabular}{l S S S S S S}
    \toprule
    Element & {$^{238}$U} & {Fe} & {Cr} & {Ni} & {O} & {F} \\
    \midrule
    Composition [kg] & 142.02 & 0.75 & 0.13 & 0.14 & 2.27 & 3104.22 \\
    \bottomrule
  \end{tabular}
  \label{table:67-salt-composition}
\end{table}

All neutronics simulations ran on OpenMC under continuous energy mode and on Moltres with the
hybrid $S_N$-diffusion method. The hybrid simulations ran with a maximum stabilization factor of
$c=250$, a void constant $\varsigma=0.5$, and a uncorrected diffusion coefficient cap of $D_g=2.5$.
Refer to Sections \ref{sec:void-treatment} and \ref{sec:diffcoef-cap} for the relevant discussion
on void treatment for near-void regions and the impact of large diffusion coefficient values on
solver performance. As for the standard neutron diffusion method, it either converges very slowly
or does not converge at all when $D_g$ values are uncapped due to excessive diffusive effects in
the air-filled control rod thimbles. Group-wise $D_g$ values, generated using OpenMC, ranged from
100 cm to 850 cm in the air regions as compared to the typical 0.5 cm to 2.0 cm range observed in
non-air regions. Therefore, all neutron diffusion method simulation results in the following
section were measured with the diffusion coefficient capped at $D_g=2.5$.

Unlike the 2-D hybrid simulations which ran with $S_8$ angular discretization, the 3-D hybrid
simulations ran with $S_6$ angular discretization due to significantly
larger computational and memory requirements demanded by the 3-D simulations. Further discussion on
computational performance may be found at the end of Section \ref{sec:3d-nts}. All simulations,
except those for temperature reactivity coefficient calculations, ran on isothermal 3-D models at
911 K.

\subsection{3-D Numerical Results \& Discussion} \label{sec:3d-nts}

\subsubsection{Initial Criticality Configuration}

The \gls{MSRE} achieved initial criticality with Rod 1 inserted to a height of 46.6 inches relative
to the fully inserted position. For reference, the rods are at a height of 51 inches when fully
withdrawn. Table \ref{table:initial-crit} shows the $k_\text{eff}$ values from \gls{MSRE}
experimental data, numerical benchmark data \cite{fratoni_molten_2020}, and this work. All
numerical models overestimate the $k_\text{eff}$ relative to the experimental value by 1-2\%
(1000-2000 pcm). In the \gls{MSRE} numerical benchmark report, the authors
attribute this $k_\text{eff}$ discrepancy largely to discrepancies in the carbon cross section data
in nuclear data libraries. They noted similar $k_\text{eff}$ overestimations by 1-2\% in benchmark
results for the HTTR, a graphite-moderated, helium-cooled reactor. The OpenMC and Moltres
simulations in this work used the same
ENDF/B-VII.1 nuclear data library as the \gls{MSRE} numerical benchmark report.

\begin{table}[htb]
  \centering
  \caption{$k_\text{eff}$ values from \gls{MSRE} experimental data, the \gls{MSRE} numerical
  benchmark \cite{fratoni_molten_2020}, and the OpenMC and Moltres models in this work.}
  \begin{tabular}{l S[table-format=1.5(2)]}
    \toprule
     & {$k_\text{eff}$} \\
     \cmidrule(l){2-2}
    \gls{MSRE} experimental data & 1.00000(420) \\
    Serpent 2 (Numerical benchmark) & 1.02132(3) \\
    OpenMC (This work) & 1.01308(20) \\
    Hybrid (This work) & 1.01955 \\
    Diffusion (This work) & 1.01885 \\
    \bottomrule
  \end{tabular}
  \label{table:initial-crit}
\end{table}

Of the four $k_\text{eff}$ values, we expect to see the closest agreement between the Serpent 2
model from the numerical benchmark and OpenMC from this work. However, the OpenMC \gls{MSRE} model
differs from the Serpent 2 \gls{MSRE} models due to geometry modifications listed in Sections
\ref{sec:2d-model-setup} and \ref{sec:3d-model-setup}. The benchmark report quantified the impact
of several geometry modifications, namely the removal of the flow distributor, the homogenization
of the sample basket, and the removal of the thermal shield. These modifications contribute to
-$0.098\%$, -$0.037\%$, and -$0.885\%$ change in the $k_\text{eff}$, respectively, for a final
$k_\text{eff}$ estimate of $1.01093(3)$. This brings the OpenMC $k_\text{eff}$ estimate of
$1.01308(20)$ to closer agreement with the numerical benchmark value.

The hybrid method $k_\text{eff}$ estimate is 0.00647 higher than the OpenMC estimate. For this
initial criticality configuration, the neutron diffusion method $k_\text{eff}$ estimate is closer
to OpenMC than the hybrid method. Overall, the OpenMC and Moltres models show good agreement with
the Serpent 2 model from the numerical benchmark.
Figures * show the neutron group fluxes on the interior of the 3-D models. The salt-graphite
lattice structure is visible from the group flux distributions. Group 7 and 8 neutron fluxes are
particularly small near the control rods due to strong thermal neutron capture cross section of
gadolinium.

\subsubsection{Control Rod Worth}

After achieving initial criticality, \gls{MSRE} researchers ran control rod calibration experiments
with the \gls{MSRE} at zero-power (typically at about 10 W power output)
\cite{prince_zero-power_1968}. They measured differential rod worths by measuring the time taken
for the reactor power to increase by 20 W following a prescribed withdrawal of Rod 1 and applying
the inhour equation to obtain the reactivity inserted. They repeated this process at different rod
heights after successive additions of highly enriched uranium capsules. The final integral rod
worth curve was normalized to the critical $^{235}$U loading of 65.25 kg
\cite{prince_zero-power_1968}.

For control rod worth calculations, the OpenMC and Moltres simulations ran with Rod 1 at fourteen
different rod positions from full insertion (0 inch) to full withdrawal (51 inches). The rod
positional limits of 0 inch and 51 inches represents the full range of deliberate rod motion
allowable by the control rod drive motor \cite{robertson_msre_1965}.

Figure \ref{fig:rod-worth} shows the integral rod worth curves from the \gls{MSRE} experimental
data and the OpenMC and Moltres numerical models. By taking the $k_\text{eff}$ when Rod 1 is fully
inserted as reference, all rod worth curves start at 0 pcm. The change in integral rod worth,
represented by the gradient in the plot, is steepest as the tip of the rod passes through the
center of the core. The hybrid method performs well in reproducing the rod worth curve from OpenMC;
all hybrid method rod worth values fall within one or two standard deviations of OpenMC.

Due to the experimental procedure of computing integral rod worth through cumulative differential
rod worth measurements, the uncertainty range of the \gls{MSRE} data scales with rod position.
Both OpenMC and the hybrid method report consistently larger integral rod worths than the
\gls{MSRE} data across all rod positions. Nevertheless, they are in good agreement with the
\gls{MSRE} data as compared to the neutron diffusion method, which significantly overestimates the
integral rod worth.
Table \ref{table:rod-worth} shows the total rod worth of Rod 1. Similar to observations from Figure
\ref{fig:rod-worth}, OpenMC and the hybrid method report total rod worths marginally larger than
the \gls{MSRE} data by 5.1\% and 4.2\%, respectively, while the diffusion method exceeds the
experimental value by 21.1\%.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\columnwidth]{rod-worth-curve}
  \caption{Reactivity inserted by Rod 1 at various rod positions relative to the full insertion.}
  \label{fig:rod-worth}
\end{figure}

\begin{table}[t]
  \centering
  \caption{Total rod worth of Rod 1 when fully inserted.}
  \begin{tabular}{l S[table-format=4(2)] S}
    \toprule
     & {$\rho_\text{worth}$ [pcm]} & {Percentage discrepancy [\%]}\\
     \cmidrule(rl){2-2} \cmidrule(l){3-3}
    \gls{MSRE} experimental data & 2250(67) & {-}\\
    OpenMC (This work) & 2364(44) & 5.1 \\
    Hybrid (This work) & 2345 & 4.2 \\
    Diffusion (This work) & 2725 & 21.1 \\
    \bottomrule
  \end{tabular}
  \label{table:rod-worth}
\end{table}

The most likely factor for OpenMC and the hybrid method overpredicting the control rod worth is the
modeling of control rods as solid annular cylinders as opposed to 36 separate control elements
strung along a flexible stainless steel cable. Figure \ref{fig:rod-bead} depicts a poison element
and its physical dimensions. Other than the Gd$_2$O$_3$-Al$_2$O$_3$ poison material, each control
element has a top and bottom closure made of INOR-8 alloy. According to the physical dimensions,
the closures make up 15.1\% of the total element height. The 3-D numerical \gls{MSRE} model in this
work omits the outer and inner INOR-8 tubes and the flexible stainless cable, and replaces the top
and bottom closure with poison material. Fully resolving the closures in Moltres would require
excessively fine axial mesh resolution and advanced adaptive meshing features not currently
available. A straightforward corrective measure for future \gls{MSRE} simulations would be to
retain the solid annular cylinder structure and adjust the rod dimensions or the rod material
composition to match the expected worth from \gls{MSRE} data or OpenMC simulations with the beaded
geometry.

\begin{figure}[t]
  \begin{subfigure}[b]{0.33\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{rod-bead-2}
  \end{subfigure}
  \begin{subfigure}[b]{0.65\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{rod-bead}
  \end{subfigure}
  \caption{Cutaway view and dimensions of a control rod poison element. Retrieved from
  \cite{tolson_msre_1967} and \cite{robertson_msre_1965}.}
  \label{fig:rod-bead}
\end{figure}

Overall, the results show that the hybrid $S_N$-diffusion method accurately reproduces the control
rod worth at various levels of insertion. As observed in the 1-D and 2-D models, the 3-D model
$k_\text{eff}$ values exhibit some discrepancies relative to OpenMC, but the discrepancies do not
affect rod worth estimates because they are systematic errors. This \gls{MSRE} hybrid model may be
improved in future work by characterizing and applying neutron albedo at the vacuum boundaries.

\subsubsection{Temperature Reactivity Coefficient}

This work compares isothermal temperature reactivity coefficients calculated from $k_\text{eff}$
values measured at 872 K and 972 K with 67.86 $^{235}$U loading. This approach is identical to that
of the \gls{MSRE} experimental measurements \cite{prince_zero-power_1968} and the \gls{MSRE}
numerical benchmark calculations \cite{fratoni_molten_2020}. The temperature reactivity coefficient
is calculated by dividing the difference in reactivities by the temperature change (100 K). All
simulations included the effects of thermal expansion on the material density but left all reactor
physical dimensions unchanged.

Table \ref{table:temp-coef} compares the isothermal reactivity coefficients from the experimental
and numerical sources. The OpenMC and Moltres models in this work show closer agreement with the
\gls{MSRE} data than the Serpent 2 model from the \gls{MSRE} numerical benchmark. All numerical
results fall within one standard deviation of the \gls{MSRE} data.

\begin{table}[t]
  \small
  \centering
  \caption{Isothermal temperature reactivity coefficients with 67.86 kg $^{235}$U loading from
  \gls{MSRE} data, the \gls{MSRE} numerical benchmark \cite{fratoni_molten_2020}, and the OpenMC
  and Moltres models in this work.}
  \begin{tabular}{l S[separate-uncertainty=true,table-format=2.2(2)] S}
    \toprule
     & {Temperature coefficient [pcm K$^{-1}$]} & {Percentage discrepancy [\%]}\\
     \cmidrule(rl){2-2} \cmidrule(l){3-3}
    \gls{MSRE} experimental data & 13.41(150) & {-}\\
    Serpent 2 (Numerical benchmark) & 12.51(23) & -6.7 \\
    OpenMC (This work) & 13.85(43) & 3.3 \\
    Hybrid (This work) & 13.98 & 4.3 \\
    Diffusion (This work) & 14.01 & 4.5 \\
    \bottomrule
  \end{tabular}
  \label{table:temp-coef}
\end{table}

\subsubsection{Computational Performance}

Beyond demonstrating the hybrid $S_N$-diffusion method, its computational performance is of
interest as a candidate for 3-D, time-dependent reactor simulations. All simulations ran on the
Polaris supercomputer system at \gls{ANL}. Each compute node consists of a 2.8-GHz AMD EPYC Milan
7543P 32-core CPU, four NVIDIA A100 GPUs, and 512 GB of RAM. The GPUs were not utilized in this
work. Moltres uses the \gls{MPI} communication protocol for parallelizing simulations. For this
work, simulations ran with 1-to-1 ratio of \gls{MPI} ranks and processor cores (i.e., 32 \gls{MPI}
ranks for each compute node).

Every 3-D hybrid method $k$-eigenvalue
simulation of the \gls{MSRE} consisted of 747,688,760 \glspl{DOF} and 1,690,777,728 \glspl{DOF} for
the neutron diffusion and $S_N$ components, respectively. Each \gls{DOF} corresponds to a scalar
or angular group flux variable value defined at a mesh node. Each simulation ran with the distributed
mesh feature, which partitions and distributes the mesh across however many \gls{MPI} ranks used
in the simulation. All variable and mesh-linked data are also distributed with the mesh. This
greatly reduces memory requirements compared to storing the full mesh for each \gls{MPI} rank.

Each 3-D \gls{MSRE} simulation required at least forty nodes for sufficient memory to run. This
sets a strict minimum requirement for the 3-D hybrid simulations. On forty nodes, the simulations
required approximately 2.2 h to complete for a total of 88 node-hours or 2816 core-hours. The
distributed mesh capability comes with the cost of increased data transfer times between the $S_N$
and diffusion solvers due to increased inter-core data transfers. Transfers accounted for
approximately 0.675 h or 30 \% of the total simulation time. Future research into optimizing mesh
distribution and transfer-related data caching may help to minimize data transfers and speed-up
hybrid method simulations. The solution time for the initial neutron diffusion calculation with zero
drift correction also takes up approximately 30 \% of the total simulation time. After subtracting
the minor computational cost of evaluating the drift term even if it is zero, the hybrid method
requires about four times the solution time of the standard neutron diffusion method.

For the hybrid method to be viable for large time-dependent simulations, it must scale reasonably
well on computing clusters. For this work, I investigated strong scaling, which measures how the
solution time decreases when more cores are used while keeping the problem size fixed. This strong
scaling study used 3-D quarter-core models run on 10, 20, 30, and 40 compute nodes.
Figure * shows the results of this scaling study.
The hybrid method implementation scales well as shown by the nearly linear scaling between the
solution time and the number of processors. Therefore, the hybrid method appears promising for the
modeling of larger reactor analyses with little scaling performance penalty.

% Strong scaling tests & discussion
% Difficulties with 3-D modeling
% SN void treatment constants, Transfer times

@@

This chapter covers work demonstrating the hybrid $S_N$-diffusion method implemented in Moltres
through 1-D, 2-D, and 3-D neutronics simulations of \gls{MSRE} models. The hybrid method combines
the $S_N$
and diffusion methods by applying transport corrections generated using the high-fidelity $S_N$
method near control rods while treating the rest of the problem with the lower-fidelity diffusion
method for tractable 3-D time-dependent reactor analyses. This is achieved by identifying and
setting up a subregion containing the control rod and its vicinity for $S_N$ transport
calculations. The $S_N$ and diffusion solvers are coupled through boundary sources at the subregion
boundaries. The hybrid method applies an adaptive algorithm for discarding inaccurate transport
correction values near the subregion boundaries and preserving the smoothness of the neutron flux
gradient. In this chapter, the 1-D simulations provided verification of the hybrid method for
estimating control rod worth. The simulations also compared two different implementations of the
hybrid method which differed by the use of drift or diffusion correction schemes.
The 2-D and 3-D simulations extended the hybrid method to multi-dimensional problems. Additionally,
the 3-D simulations validated the numerical \gls{MSRE} model against \gls{MSRE} experimental data
and provided useful insight on the computational performance of the hybrid method for
large reactor analyses.

Six 1-D test cases based on the \gls{MSRE} with increasing geometric complexity were run using the
standard $S_N$ method, the hybrid method, and standard neutron diffusion method and verified against
reference OpenMC Monte Carlo neutron transport simulations under continuous neutron energy (OpenMC-CE) and
multigroup (OpenMC-MG) modes. The $S_N$ implementation with nonlinear diffusion acceleration
showed good agreement with OpenMC-MG which indicated that discrepancies between those codes and
OpenMC-CE largely arose from the eight neutron energy group structure for multigroup simulations.
Both hybrid method implementations, with drift or diffusion correction schemes, showed similar
performance in the 1-D cases. While the $k_\text{eff}$ estimates from the hybrid method deviated
from the $S_N$ method and OpenMC-MG by 100 to 400 pcm, the hybrid method accurately reproduces
control rod worths. Rod worth estimates from OpenMC-MG, $S_N$, and hybrid methods differed from
OpenMC-CE by 2 to 3 \%. The hybrid method improves on the neutron diffusion method whose rod worth
error estimate is 8 \%. The hybrid method also produced better agreement in neutron flux
distributions with the neutron transport methods compared to the neutron diffusion method.

Analysis on the drift and diffusion correction schemes showed that the diffusion correction
parameter led to numerous discontinuities tending to infinity values near points in the problem
domain where flux peaks and troughs occur. These discontinuities and negative diffusion coefficient
values are unphysical and would lead to numerical stability issues. Therefore, the hybrid method
implementation with drift correction scheme was chosen for the remaining 1-D analyses and the
2-D and 3-D analyses.

Analysis on the impact of $S_N$ correction subregion size found that it has little impact on the
control rod worth estimates as long as the subregion size is kept consistent (i.e., same size) when
performing control rod worth measurements. Rod worths varied by less than 0.35 \% across the
different subregion sizes investigated. A separate analysis showed that the $k_\text{eff}$
estimates from the hybrid method exhibit superlinear convergence rate when tightening the $S_N$
convergence tolerance value. The convergence tolerance value also impacts the number of outer
iterations between $S_N$ and diffusion solvers. After just two outer iterations, the error in
$k_\text{eff}$ fell to below 10$^{-7}$.

The hybrid method continued to show improved control rod estimates over the neutron diffusion
method in 2-D quarter-core and full-core \gls{MSRE} simulations with up to three rods inserted in the
full-core model. The hybrid method reported rod worth error magnitudes of less than 40 pcm relative
to OpenMC-CE, which are significant improvements over the neutron diffusion method error estimates
ranging from 569 pcm to 1484 pcm. The hybrid method also showed significant improvements in the
absolute mean and maximum errors in fuel channel power distributions relative to the reference
OpenMC-CE distribution.

For 3-D simulations, the \gls{MSRE} experimental data and the \gls{MSRE} numerical benchmark
report \cite{fratoni_molten_2020} both served as references for model validation. The numerical
benchmark is included in the International Reactor Physics Experiment Evaluation Project (IRPhEP)
handbook. The
$k_\text{eff}$ estimates of the \gls{MSRE} at initial criticality from OpenMC, the hybrid method,
and the neutron diffusion method in this work showed good agreement with the \gls{MSRE} numerical
benchmark. All numerical $k_\text{eff}$ estimates, including that from the benchmark report
\cite{fratoni_molten_2020}, exceeded the \gls{MSRE} experimental value by 1-2 \% due to possible
biases and uncertainties in the neutronic properties of graphite. Temperature reactivity
coefficient values from OpenMC, the hybrid method, and the neutron diffusion method also show good
agreement with the \gls{MSRE} data within experimental uncertainty with percentage discrepancies of
about 3 \%, but differ from the numerical benchmark which has a percentage discrepancy of -6.7 \%.

Control rod worth calculations were also performed for Rod 1 of the \gls{MSRE} under various levels
of insertion from its fully inserted to its fully withdrawn state. OpenMC and the hybrid method
showed excellent agreement at every stage within one standard deviation. Due to geometric
approximations of the control rods in the numerical models, they overestimated the total
rod worth relative to the \gls{MSRE} experimental data by approximately 4-5 \%. The hybrid method
compares favorably against the neutron diffusion method which overestimated the total worth by
21.1 \%.

The 3-D simulations performed well under the strong scaling test which studies how the solution
time scales with the number of processors used to run the simulation. The 3-D simulations showed
close to linear scaling within the scope of the strong scaling test. Therefore, the hybrid method
exhibits good computational scaling performance which is essential for large 3-D reactor analyses.
Excessively large memory requirements were mitigated through the distributed mesh capability in
Moltres. Some performance optimizations may be possible to reduce data transfer times
between the $S_N$ and diffusion solvers which currently take up about 30 \% of the total solution
time as a result of the inter-processor communications.

Overall, the hybrid $S_N$-diffusion method provides a significant improvement over the neutron
diffusion method at only four times the computational expense. Control rod estimates
show good agreement with reference Monte Carlo solutions and \gls{MSRE} experimental data.
Further investigations into characterizing and minimizing neutron leakage rates at vacuum
boundaries may help with reducing $k_\text{eff}$ discrepancies with reference Monte Carlo
solutions.
